---
title: "jonssvatnet microcrustacean time-series"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(mapview)
```

# Download and wrangle the Jonsvatnet zooplankton time-series from GBIF

The dataset is part of the "Liminic freswhater pelagic collection" published in the dataset [Limnic freshwater pelagic invertebrates biogeographical mapping/inventory NTNU University Museum](https://www.gbif.org/dataset/edd9b710-f3bb-4c49-9fa5-724cd4ecfc7d)

The zooplankton time-series from lake Jonsvatnet is selected based upon geograpic locations of the sampling stations. See the [download page](https://doi.org/10.15468/dl.m55nyd) for details on search parameters. 

Description of the time-series (in Norwegian) given in NTNU University Museum reports - [latest from 2019](https://www.ntnu.no/documents/10476/1291996359/2020-4+Notat+Jonsvatnet+2019.pdf/01effd35-2707-bf15-d684-1c24535dfb69?t=1601285434082)

## Download data 

```{r download, include=FALSE}

download_url <- "https://api.gbif.org/v1/occurrence/download/request/0239745-200613084148143.zip"

temp <- tempfile()
download.file(url=download_url, destfile=temp,quiet=FALSE)
data <- read.table(unzip(temp, "occurrence.txt"),sep="\t",header=T)

```

# Preliminary data investigation 
Have a look at the data broken down on sampling stations, dates, number of sampling events and sampling protocol. 

```{r data-investigation, include=FALSE}
data_overview <- data %>% 
  select(taxonKey,eventDate,fieldNumber,year,eventID,samplingProtocol,
         locality,locationID,decimalLatitude,decimalLongitude)

data_overview_table <- data_overview %>%
  

map_data <- data_overview %>%
  select(locality,locationID,decimalLatitude,decimalLongitude) %>%
  group_by(locality,locationID,decimalLatitude,decimalLongitude) %>%
  summarise()
map_data_sf = st_as_sf(map_data, coords = c("decimalLongitude", "decimalLatitude"), 
                 crs = 4326)

mapview(map_data_sf)

```
